---
title: "Snowball Searches"
vignette: >
  %\VignetteIndexEntry{test}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
format:
  html:
    toc: true
    number-sections: true
execute:
  eval: false
  cache: false
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'  
---

```{r}
#| label: setup
#| include: false

library(openalexSnowball)
```

# Overview

`openalexSnowball` provides snowball search utilities that build on the
`openalexPro` pipeline. A snowball search starts from a set of key papers and
follows citation links to collect cited and citing works. Results are stored on
disk as parquet datasets and can be loaded back into R using `read_snowball()`.

# Basic workflow

## 1. Define keypapers

Keypapers can be identified by OpenAlex IDs or DOIs.

## 2. Run a snowball search

```{r}
#| label: snowball-openalex-ids

snowball_dir <- "./snowball_ids"
pro_snowball(
  identifier = c("W2741809807", "W2755950973"),
  output = snowball_dir,
  verbose = TRUE
)
```

```{r}
#| label: snowball-dois

snowball_dir_dois <- "./snowball_dois"
pro_snowball(
  doi = c("10.1016/j.joi.2017.08.007", "10.7717/peerj.4375"),
  output = snowball_dir_dois,
  verbose = TRUE
)
```

# Reading results

Use `read_snowball()` to load nodes and edges. The output can be an Arrow
Dataset (default) or collected into memory as tibbles.

```{r}
#| label: read-snowball

snowball <- read_snowball(
  snowball = snowball_dir,
  edge_type = c("core", "extended"),
  return_data = TRUE,
  shorten_ids = TRUE
)

snowball$nodes
snowball$edges
```

# Output layout

A snowball search writes a folder structure similar to the following:

```
snowball/
├── cited_json/
├── cited_jsonl/
├── citing_json/
├── citing_jsonl/
├── edges/
├── keypaper_json/
├── keypaper_jsonl/
├── keypaper_parquet/
└── nodes/
```

# Tips

- Use `return_data = FALSE` to keep large datasets on disk.
- Use `edge_type = "core"` when you only need edges involving keypapers.
